Security Properties for Stack Safety

### 2021-05-13 short/main.tex
Polishing / comments

The call stack is a perennial target for low-level attacks, leading to
consequences ranging from leakage or corruption of private stack data to
control-flow hijacking. To prevent or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}.
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent.
At best, they define stack safety by reference to an idealized machine that
is arguably stack safe by construction \citep{SkorstengaardSTK}.
But these mechanisms can be intricate, and it would be useful to have a precise, generic, and formal
specification for stack safety, both to compare the security claims of different
enforcement techniques and to rigorously validate such claims.

We propose such a characterization,
using the technical framework of language-based security.
Stack safety protects a caller from its callee, guaranteeing
the {\em integrity} and {\em confidentiality} of the caller's local state
until it regains control. This formulation not only captures the intuition
that the callee cannot directly access the caller's state, but gives a novel
way of looking at control-flow attacks such as return-oriented-programming (ROP).
In an ROP attack, the callee pretends to return to its caller, thus accessing the
caller's privileges, but does so in such a way that it is still in control of
execution. Our model allows the caller to dictate the terms of a valid return --
typically, that the stack pointer is restored and execution proceeds from the
return address. An ROP attack that ``returns'' and reads the caller's state is
therefore treated precisely as if the callee had never returned, but attempted
to read the caller's state directly. It is a confidentiality violation. Likewise
an attack that would corrupt the caller's state is an integrity violation.

We formalize integrity and confidentiality as trace properties, in two different
variants: {\em stepwise} variants, in which a caller's data is {\em never}
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional, and
any reasonable protection mechanism ought to enforce them,
even if it does not prevent every single dangerous read or write.
Confidentiality is especially interesting, as it is based on the traditional
notion of noninterference. But where noninterference is normally presented as an
end-to-end hyperproperty, stack confidentiality is noninterference applied over
multiple nested subtraces -- one for every call.

To demonstrate the utility of our properties, we use them
to validate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that these micro-policies correctly abort programs that
would violate stack safety.

>>>>

The call stack is a perennial target for low-level attacks, leading to
consequences ranging from leakage or corruption of private stack data to
control-flow hijacking. A profusion of
software and hardware protections have been proposed for detecting or
preventing such attacks,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}.
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent.
At best, they define stack safety by reference to an idealized machine that
is arguably stack safe by construction \citep{SkorstengaardSTK}.
But these mechanisms can be intricate, and it would be useful to have a precise, generic, and formal
specification for stack safety, both to compare the security claims of different
enforcement techniques and to rigorously validate such claims.

We propose such a characterization,
using the technical framework of language-based security.
Stack safety protects a caller from its callee, guaranteeing
the {\em integrity} and {\em confidentiality} of the caller's local state
until control returns to it. This formulation not only captures the intuition
that the callee cannot directly access the caller's state, but gives a novel
perspective on control-flow attacks such as return-oriented-programming (ROP).
In an ROP attack, the callee pretends to return to its caller, thus accessing the
caller's privileges, but does so in such a way that it is still in control of
execution\bcp{that's not very clear to me}. Our model allows the caller to
dictate the terms of a valid return -- 
typically, that the stack pointer is restored and execution proceeds from the
return address. An ROP attack that ``returns'' and reads the caller's state is
therefore treated precisely as if the callee had never returned, but attempted
to read the caller's state directly: it is a confidentiality
violation. Likewise, an attack that would corrupt the caller's state is an
integrity violation.

We formalize integrity and confidentiality as trace properties\bcp{hyperproperties??}, in two different
variants: {\em stepwise} variants, in which a caller's data must {\em never} be
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional, and
any reasonable protection mechanism ought to enforce them,
even if it does not prevent every single dangerous read or write.
Confidentiality is especially interesting, as it is based on the traditional
notion of noninterference. But where noninterference is normally presented as an
end-to-end hyperproperty, stack confidentiality is noninterference applied over
multiple nested subtraces -- one for every call.

To demonstrate the utility of our properties, we use them
to validate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Rossleer and Dehon's micro-policies correctly abort programs that
would violate stack safety.
